---
title: "Calculating attributable numbers and rates in `cityHeatHealth`"
output:
  rmarkdown::html_vignette:
    fig.retina: 2
    dev: svg
vignette: >
  %\VignetteIndexEntry{Calculating attributable numbers and rates in `cityHeatHealth`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cityHeatHealth)
```

Estimating Attributable numbers (and rates of attributable numbers) are a key way that we can translate relative risks into numbers that are more tangible in public health settings. Below we provide easy functionality to go from our model objects to estimates of attributable numbers and rates.

### Setup

The first step of calculating attributable numbers is having a population data estimate. 

This varies a lot by place and dataset, so we don't include functionality for it (but an example of how this could be done can be seen in `vignette("get_pop_estimates")`).

Assume you are starting with a dataset for the entire timeframe that looks like this:
```{r}
library(data.table)
data("ma_pop_data")
setDT(ma_pop_data)
ma_pop_data
```

Need to do some transformations:

* pivot longer
* variable clean

Note again, this processing will vary by application so this approach is not prescriptive !

Pivot longer:
```{r transform}
ma_pop_data_long <- melt(
  ma_pop_data,
  id.vars = "TOWN20",
  variable.name = "sex_age",
  value.name = "population"
)
```

Variable clean:
```{r clean}
ma_pop_data_long$sex_age <- as.character(ma_pop_data_long$sex_age)
varnames <- strsplit(ma_pop_data_long$sex_age, "_", fixed = T)
varnames <- data.frame(do.call(rbind, varnames))
names(varnames) <- c('sex', 'age_grp')
rr <- which(varnames$sex == 'Female')
varnames$sex[rr] <- 'F'
rr <- which(varnames$sex == 'Male')
varnames$sex[rr] <- 'M'
ma_pop_data_long$sex = varnames$sex
ma_pop_data_long$age_grp = varnames$age_grp
ma_pop_data_long$sex_age <- NULL
ma_pop_data_long
```

We assume that these properties hold for the entire timeframe of our analysis, but you could also make a version of this dataset with a 'year' column. 

Now, quickly get a `condPois_1stage()` and `condPois_2stage()` objects to use in testing:
exposures
```{r ee}
library(data.table)
exposure_columns <- list(
  "date" = "date",
  "exposure" = "tmax_C",
  "geo_unit" = "TOWN20",
  "geo_unit_grp" = "COUNTY20"
)

ma_exposure_matrix <- make_exposure_matrix(
  subset(ma_exposure, COUNTY20 %in% c('MIDDLESEX', 'WORCESTER') &
           year(date) %in% 2012:2015), 
  exposure_columns)
```

outcomes
```{r oo}
outcome_columns <- list(
  "date" = "date",
  "outcome" = "daily_deaths",
  "factor" = 'age_grp',
  "factor" = 'sex',
  "geo_unit" = "TOWN20",
  "geo_unit_grp" = "COUNTY20"
)

ma_outcomes_tbl <- make_outcome_table(
  subset(ma_deaths,COUNTY20 %in% c('MIDDLESEX', 'WORCESTER') &
           year(date) %in% 2012:2015), outcome_columns)

```

models
```{r jj}
ma_model <- condPois_2stage(ma_exposure_matrix, ma_outcomes_tbl, verbose = 1)
```

### Estimating the AN

Ok so now you pass in `population`. 

So now estimate the AN as a full object

Remember that this needs to be compatible for:

* single zone

* ma model with ma_model$`_`

* ma model with factor ma_model$`0-17`
  > I think you can handle this the same way you did before, with recursion
  

Now in this second step, you can choose the aggregation level that you want results to.

In this block you need:
  
  * what spatial resolution are you summarizing to:
  ->> 'geo_unit', 'geo_unit_grp', or 'all'

* are you just getting the impacts that are > then the centering point: 
  ->> lets just assume yes for now, can always go back and change it

```{r calcAN}
ma_AN <- calc_AN(ma_model, ma_outcomes_tbl, ma_pop_data_long,
                 agg_type = 'TOWN20', 
                 join_cols = 'TOWN20', 
                 nsim = 100,
                 verbose = 2)
ma_AN$`_`$rate_table
ma_AN$`_`$number_table
```

you can change `agg_type` to be a different spatial resolution -- either whatever the group variable was or "all"
```{r grpsum}
ma_AN <- calc_AN(ma_model, ma_outcomes_tbl, ma_pop_data_long,
                 agg_type = 'COUNTY20', 
                 join_cols = 'TOWN20', 
                 nsim = 100,
                 verbose = 2)
ma_AN$`_`$rate_table
ma_AN$`_`$number_table
```

See that the numbers are roughly the same for Suffolk county ? They won't be exactly the same because of how the averaging works.

Some plot functions exist:
```{r grpsum3, fig.height=3, fig.width=8}
plot(ma_AN, table_type = 'rate', above_MMT = T)
plot(ma_AN, table_type = 'rate', above_MMT = F)
```

### Estimating the AN - single

check of single
```{r checkSingle1, fig.height=3, fig.width=8}

# run the model
m2 <- condPois_1stage(exposure_matrix = ma_exposure_matrix, 
                  outcomes_tbl = ma_outcomes_tbl, 
                  multi_zone = TRUE)

ma_AN_s1 <- calc_AN(m2, ma_outcomes_tbl, ma_pop_data_long,
                 agg_type = 'COUNTY20', 
                 join_cols = 'TOWN20', 
                 nsim = 100,
                 verbose = 2)

ma_AN_s1$`_`$rate_table
plot(ma_AN_s1, "num", above_MMT = T)

```

### Estimating the AN - with factors

In the case where you have factors, you can easily extend this

```{r fctrun, fig.height=3, fig.width=8}
ma_outcomes_tbl_fct <- make_outcome_table(
  subset(ma_deaths,COUNTY20 %in% c('MIDDLESEX', 'WORCESTER') &
           year(date) %in% 2012:2015), 
  outcome_columns, collapse_to = 'age_grp')

ma_model_fct <- condPois_2stage(ma_exposure_matrix, ma_outcomes_tbl_fct, verbose = 1)

ma_AN_fct <- calc_AN(ma_model_fct, ma_outcomes_tbl_fct,
                     ma_pop_data_long,
                 agg_type = 'COUNTY20', 
                 join_cols = 'TOWN20', 
                 nsim = 100,
                 verbose = 1)

plot(ma_AN_fct, "num", above_MMT = T)

```

These results are fictional of course but show what kind of outputs can be made
easily.

```{r multi_plot3d, fig.height=10, fig.width=6}
spatial_plot(ma_AN_fct, shp = ma_counties, table_type = "num", above_MMT = T)
```
