---
title: "Quickstart guide to `cityHeatHealth`"
output:
  rmarkdown::html_vignette:
    fig.retina: 2
    dev: svg
vignette: >
  %\VignetteIndexEntry{Quickstart guide to `cityHeatHealth`}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(cityHeatHealth)
```

The package `cityHeatHealth` makes it simple to estimate heat-health impacts at small spatial scales. Starting from a messy exposure and outcome dataset, we can quickly estimate heat-health impacts.

This package can be used in three main ways:

| Single-Zone | 2-stage design | Spatial Bayes |
|-------------|----------------|---------------|
| A standard single-stage conditional poisson model when estimating heat-health impacts in a single zone | A 2-stage design is used when estimating heat-health impacts across many zones, but where individual zone impacts are desired  | If numbers are very small in the 2-stage design, spatial bayesian methods can be used to tighten confidence intervals |

The vignettes for these are:

* `vignette("single_zone_demo")`
* `vignette("two_stage_demo")`
* `vignette("bayesian_demo")`

Following any of these implementations, the same attributable number calculation is applied to model outcomes to get heat-attributable outcomes. To avoid repetition, we provide a separate vignette that details the attributable number module: `vignette("attributable_number")`.

## Starting a new analysis

To start a new analysis, you will need the following **4** datasets:

| Exposure | Outcomes | Populations | Spatial |
|----------|----------|----------|----------|
| Exposures at the daily scale for each `geo_unit` | Health outcomes at the daily scale for each `geo_unit` | Population data for each subdivision of the health outcome data that you want results for | A map showing how the various `geo_unit`s are neighbors |

This package comes pre-loaded with sample datasets of each type (e.g., `ma_exposure`, `ma_deaths`, `ma_pop_data`, and `ma_towns` respectively) so each of these methods can be explored. See the above vignettes for starter code for each type of analysis.
